Unification can be generally thought of the study of formal solving of equations between terms.

This topic was introduced by Herbrand, but became really widespread after the work of J.~A.~Robinson on automated theorem proving.
The unification technique is also at the core of the logic programming language \textsc{Prolog} and type inference for functional programming languages such as \textsc{CaML} and \textsc{Haskell}.

%\subsubsection*{Terms and substitutions}

Specifically, we will be interested in the following problem: 
\begin{center}
	\it Given two terms, can they be \enquote{made equal} by replacing their variables?
\end{center}
\vspace{-1mm}
% Compacte
%To make things more concrete, we set a specific set of terms for the rest of this article.
\define[terms]
	{We consider the following set of first-order terms
% Compacte
	\[\terms::=\ \ x,y,z,\:\dots\ |\ \TT a,\tt b,\tt c,\:\dots\ |\ \terms\p\terms\]
% Compacte	
	where  $\,x,y,z,\:\dots\in \vars\,$ are variables and $\,\TT a,\TT b,\TT c,\:\dots\,$ are constants, while $\,\p\,$ is a binary function symbol.

\smallskip
For any $\,t\in\terms\,$, we will write $\,\var(t)\,$ the set of variables occurring in $\,t\,$.
We say that a term is closed when $\,\var(t)=\varnothing\,$, and denote $\,\closed\,$ the set of closed terms. 
}

\notation{We will write the binary function symbol as \emph{right associating} to avoid an excess of parenthesis: $\:t\p u\p v \::=\:t\p(u\p v) \,$}
%We also leave the $\,\p\,$ implicit $\,\TT c\p u\,$, so that: $\,\TT{lrc}\p\TT{llc}:=(\TT l \p\TT r \p\TT c)\p \TT l \p\TT l \p\TT c=(\TT l \p(\TT r \p\TT c))\p (\TT l \p(\TT l \p\TT c))\,$}

%Variables in terms are meant to be replaced when needed by a term of the language. This is implemented by the notion of substitution.

\define[substitution]
	{A substitution is a map $\,\theta:\:\vars \rightarrow \terms\,$ such that the set $\,\dom(\theta):=\{\:v\in \vars\:|\:\theta(v)\not=v\:\}\,$ (the \emph{domain} of $\,\theta\,$) is finite.
	A substitution with domain $\,\{\,x_1,\dots,x_n\,\}\,$ such that $\,\theta(x_1)=u_1\,,\,\dots\,,\,\theta(x_n)=u_n\,$ will be written as $\,\{\:x_1\mapsto u_1\,;\,\dots\,;\, x_n\mapsto u_n\:\}\,$.

\smallskip
	If $\,t\in \terms\,$ is a term we write $\,t.\theta\,$ the term $\,t\,$ where any occurrence of any variable $\,x\,$ has been replaced by $\,\theta(x)\,$.
	
	\smallskip
	If $\,\theta=\{\:x_i\mapsto u_i\:\}$ and $\,\psi=\{\:y_j\mapsto v_j\:\}\,$, their \emph{composition} is defined as
\[\theta;\psi\::=\: 
	\{\:x_i\mapsto u_i.\psi\:\} \:\cup\:
	\{\:y_i\mapsto v_i \:|\;y_i \not\in \dom(\theta)\:\}\]
\vspace{-7mm}
}

\remark{The composition of substitutions is such that $\,t.(\theta;\psi)=(t.\theta).\psi\,$ holds.}

\define[vocabulary]{A \emph{renaming} is a substitution $\,\alpha\,$ such that $\,\alpha(\vars)\subseteq \vars\,$ and that is bijective.

\smallskip Two substitutions $\,\theta,\psi\,$ are equal \emph{up to renaming} if there exist a renaming $\,\alpha\,$ such that $\,\psi=\theta;\alpha\,$.
Two terms $\,t,u\,$ are equal \emph{up to renaming} if there exist a renaming $\,\alpha\,$ such that $\,t=u.\alpha\,$.

\smallskip A substitution $\,\psi\,$ is an \emph{instance} of $\,\theta\,$ if there exists a substitution $\,\sigma\,$ such that $\,\psi=\theta;\sigma\,$.

%\smallskip A substitution $\,\theta\,$ is \emph{idempotent} if $\,\theta;\theta=\theta\,$.
}

\theorem%[folklore]
	{\label{folklore}The following properties hold for any substitutions $\,\theta,\psi\,$:
	\begin{itemize}%[parsep=0pt]
		\item The only invertible substitutions are renamings.
%		\item Every substitution is equal up to renaming to an idempotent substitution.
		\item If $\,\theta\,$ is an instance of $\,\psi\,$ and $\,\psi\,$ is an instance of $\,\theta\,$, then they are equal up to renaming.
	\end{itemize}
}

%\subsubsection*{Unification problems}

To allow easier manipulation, the problem of unifying \emph{two} terms needs to be generalized into the problem of simultaneously unifying several pairs of terms.

\define[unification]
	{A \emph{unification problem} is finite set of pair of terms $\,P=\{\:t_i\uequ u_i\:\}\,$.
	
	It is said to be unifiable if it has a \emph{unifier}: a substitution $\,\theta\,$ such that
	\[u_i.\theta=t_i.\theta\ \text{ for all }\,i\]
		In particular, we say that two terms $\,t,u\,$ are \emph{unifiable} if there exists a substitution $\,\theta\,$ such that $\,t.\theta=u.\theta\,$.
		
	We say that $\theta\,$ is a \emph{most general unifier (MGU) of $\,P\,$} if any other unifier of $\,P\,$ is an instance of $\,\theta\,$.
}
%\vspace{-5mm}

\remark{It follows from Theorem \ref{folklore} that any two MGU of the same unification problem are equal up to renaming.}

\smallskip
We will be interested mostly in the weaker variant of unification where one can first rename terms so that they variables are distinct, we introduce therefore a specific vocabulary for it.

\define[disjointness and matching]
	{\label{disjoint}Two terms $\,t,u\,$ are \emph{matchable} if $\,t',u'\,$ are unifiable, where $\,t',u'\,$ are renamings of $\,t,u\,$ such that $\,\var(t')\cap\var(u')=\varnothing\,$.
	
	If two terms are not matchable, they are said to be \emph{disjoint}.
}

\example{$x\,$ and $\,\TT f\p x\,$ are not unifiable.
However they are matchable, as $\,x.\{\,x\mapsto y\,;\,y\mapsto x\,\}=y\,$ which is unifiable with $\,\TT f\p x\,$.

More generally, disjointness is stronger than non-unifyability.}

\medskip
The crucial feature of first-order unification is the (decidable) existence of most general unifiers for unification problems that have a solution.

\theorem[MGU]
{If a unification problem has a unifier, then it has a MGU.

Whether two terms are unifiable and, in case they are, finding a MGU is decidable.}


Let us note the following fact, that will be useful in the next section: solving a unification problem can be done incrementally.

\smallskip
\notation{If $\,P=\{\:t_i\uequ u_i\:\}\,$ is a unification problem and $\,\theta\,$ a substitution, we write $\,P.\theta:=\{\:t_i.\theta\,\uequ\, u_i.\theta\:\}\,$.}

\lemma[partial unification]
	{\label{part-unif}Let $\,P=Q\uplus R\,$ ($\,\uplus\,$ denotes disjoint union) be a unification problem.
	The following statements are equivalent:
	\begin{itemize}
		\item $\,P\,$ is unifiable
		\item $\,Q\,$ is unifiable with MGU $\,\theta\,$ and $\,R.\theta\,$ is unifiable with MGU $\,\psi\,$
	\end{itemize}
In that case, we have moreover that $\,\theta;\psi\,$ is a MGU of $\,P\,$.
}

%\proof{The algorithm from the above proof can be modified to treat first pairs of $\,Q\,$ at line 3 without affecting its behaviour. 
%}

%\subsubsection*{Complexity of unification}

%Herbrand's original procedure for solving unification was non-deterministic.

%J.~A.~Robinson was the first to prove the existence of MGU for two unifiable terms and derived from this a deterministic unification procedure.
%His original procedure was however quite inefficient, with potential exponential blowups in some cases. 

%\smallskip
%It turns out that the general unification problem can be solved in linear time, which was discovered independently in \cite{paterson_linear_1978} and \cite{martelli_unification_1976}, improving on the almost-linear algorithms thus far designed.

%\smallskip
%Unification was first thought to be a \textsc{NLogspace}-complete problem. However, an error was eventually found in the note \cite{lewis_unifiability_1982} claiming this fact, which eventually led to a determination of the intrinsic complexity of the problem: unification is indeed \textsc{Ptime}-complete under logarithmic space reductions \cite{dwork_sequential_1984}.

As unification grew in importance, the study of its complexity gained in attention.
A complete survey \cite{knight_unification_1989} tells the story of the bounds getting sharpened: general first-order unification was finally proved \cite{dwork_sequential_1984} to be a \textsc{Ptime}-complete problem.

\smallskip
In this article, we are concerned with a very much simpler case of the problem: the matching (definition \ref{disjoint}) of linear terms (\textit{ie.} where variables occur at most once).
This case can be solved in a space-efficient way.

\theorem[matching in logarithmic space {\cite[Lemma 20]{dwork_parallel_1988}}]
{\label{unif-logspace}Wether two linear terms $\,t,u\,$ with disjoint sets of variables are unifiable, and if so finding a MGU, can be computed in logarithmic space in the size\footnote{The \emph{size} of a term is the total number of occurrences of symbols in it.} of $\,t, u\,$ on a deterministic Turing machine 
%\smallskip
%Moreover, their MGU is of size at most the sum of the sizes of $\,t\,$ and $\,u\,$.
}

The lemma in the article actually states that the problem is in \textsc{NC}\up 1, a complexity class of parallel computation known to be included in \textsc{Logspace}.